A Fredholm integral equation approach to computing Stokes eigenvalues in the plane

### 2019-03-12 doc/draft-01-stokes/03analysis2.tex
Second pass at section 3.6

First, we briefly review Nystr\"{o}m discretizations for integral operators.
Suppose that $\gamma(t):[0,L] \to \Gamma$ is an arc-length
paramterization of $\Gamma$, where $L$ is the length of the
curve $\Gamma$.
Using this parametrization, the integral equation
$(\cI - 2\cD_{k} - 2\cW)\bmu = \ff$, 
can be reformulated as a system of integral equations on 
$C[0,L]$ of the form 
\begin{equation}
\bmu(x) + \int_{0}^{L} \bK(x,y) \bmu(y) dy = \ff(x) \, ,  \label{inteqfreddet}
\end{equation}
where $\bK(x,y)$ is $O(|x-y|^2 \log{|x-y|})$ as $x\to y$, and
is analytic away from $x\neq y$.
Let $K$ denote the integral operator with kernel $\bK$, i.e.
\begin{equation}
K[\mu](\bx) = \int_{0}^{L} \bK(x,y) \bmu(y) dy \, .
\end{equation}

Suppose that $[0,L]$ is discretized using $N$ discretization
nodes $x_{j}^{N}$, $j=1,2,\ldots N$, along with an
the quadrature rule, $Q_{N}(x,x_{j,N})$, $j=1,2,\ldots N$, $x\in[0,L]$.
Let $K^{N}$ be the finite dimensional approximations of $K$ 
given by:
\begin{equation}
K^{N}[\phi](\bx) = \sum_{j=1}^{N} Q^{N}(x,x^{N}_{j})  
\bK(x,x_{j}^{N}) \phi(x,x_{j}^{N}) \, .
\end{equation}
We will refer to $K^{N}$ is the Nystr\"{o}m approximation of $K$ 
with quadrature rule $Q^{N}$

\begin{definition}
The Nystr\"{o}m approximation of $K$ with quadrature rule $Q^{N}$ 
is considered to be $p$th order accurate 
if there exists a $C$ independent of $N$
such that for all $\bmu \in C^{p}$, 
\begin{equation}
\lim_{N\to \infty} |(K^{N} - K)\bmu| \leq C_{\bmu} N^{-p} \, .
\end{equation}
\end{definition}

\begin{remark}
There are many sufficient conditions on
the choice of $x_{j,N}, Q_{N}(x,x_{j,N})$ $j=1,2\ldots,N$, $N \in \mathbb{N}$
for which the corresponding Nystr\"{o}m discretizations are $p$th order accurate.
A few such sets of sufficient conditions are discussed in~\cite{kress1989linear}.
While only the Kress quadrature~\cite{kress1991boundary} have been proven to be 
spectrally 
accurate, i.e.
\begin{equation}
\lim_{N\to\infty} |(K^{N}-K)\bmu| \leq C_{\bmu} e^{-\alpha N} \quad \text{for some} \, \alpha>0 \, ,
\end{equation}
several standard quadrature methods such as 
Alpert quadrature~\cite{alpert1999}, quadrature by expansion~\cite{qbxquadrature}, 
generalized Gaussian quadratures~\cite{bremer2010}
to name a few demonstrate $p$th order convergence numerically. 
See~\cite{hao2014high} for a detailed discussion. 
\end{remark}

Let $\tilde{K}^{N}$ denote the matrix  
corresponding to $K^{N}$, which is given by
\begin{equation}
\tilde{K}^{N}_{i,j} = Q^{N}(x_{i}^{N}, x_{j}^{N}) \bK(x_{i}^{N}, x_{j}^{N})
\quad i,j=1,2\ldots N \, .
\end{equation}

\begin{remark}
When solving the second kind integral equation~\cref{inteqfreddet} using the Nystr\"{o}m 
method, one solves the linear system $(I + \tilde{K}^{N})\tilde{\bmu} = \ff(x_{j}^{N})$,
for the unknown density $\tilde{\bmu} = \{ \bmu_{j} \}_{j=1}^{N}$ 
and then using the natural interpolation
\begin{equation}
\bmu(x) = \ff(x) - \sum_{j=1}^{N} Q^{N}(x,x_{j}^{N}) \bK(x,x_{j}^{N}) \bmu_{j}^{N} \, .
\end{equation}
\end{remark}

%The standard Nystr\"{o}m procedure is to solve the integral equation
%\begin{equation}
%\bmu_{i,N} + \sum_{j=1}^{N} Q_N(x_{i,N},x_{j,N} \bK(x_{i,N},x_{j,N})
%\bmu_{j,N} = \ff(\bx_{i,N}) \, ,
%\end{equation}
%for the unknowns $\bmu_{i,N}$, along with the natural interpolant
%\begin{equation}
%\bmu(\bx) = \ff(\bx) - \sum_{j=1}^{N} Q_N(x,x_{j,N}) \mu_{j,N} \, .
%\end{equation}

%\begin{equation}
%\end{equation}
%Suppose that the quadrature rule $Q_{N}$ satisfies
%the following two additional constraints:
%a) 
%\begin{equation}
%\sum_{j=1}^{N} |Q_{N}(x,x_{j,N})| \leq C \, ,
%\end{equation}
%where $C$ is a constant independent of $x$ and $N$, and b) that
%\begin{equation}
%\lim_{x\to y} \sup_{n \in \mathbb{N}} \sum_{j=1}^{N} 
%|Q_{N}(x,x_{j,N}) - Q_{N}(y,x_{j,N})| \to 0 \, .
%\end{equation}

%Suppose that the domain $[0,L]$, is partitioned into $N$ 
%panels of equal length, where each panel is discretized 
%using order $p$ discretization nodes (e.g.
%Let $\cL = -2\cD_{k} - 2\cW$, and let $\cJ_{1}(X)$ denote the 
%space of trace class operators on the Hilbert Space $X$.
%Suppose the boundary $\Gamma$ is parametrized into $N$ panels
%of equal length.

The following theorem is the main result of the section
and shows that the determinant of the
linear system corresponding to the Nystr\"{o}m 
discretization of $\cI - 2\cD_{k} - 2\cW$ 
converges to $0$ precisely at the eigenvalues of the Dirichlet
problem for Stokes equation on
with error proportional to the quadrature 
method used in the Nystr\"{o}m discretization.

\begin{thrm}
\label{thm:mainconvfreddet}
Suppose that $\Omega$ is a simply connected domain with an analytic boundary 
$\Gamma$. 
Suppose the Nystr\"{o}m method with quadrautre rule $Q^{N}$ 
is $p$th order accurate. 
Let $\tilde{L}^{N}_{k}$ denote the finite dimensional linear system corresponding
to $-2\cD_{k} -2\cW$.
Let $f(k) = \text{det}(I + \tilde{L}^{N}_{k})$. 
Then $f:[0,R] \to \mathbb{C}$  is an analytic function of $k$ 
for any $R<\infty$. 
Suppose further that $\tilde{k}$ is an Dirichlet eigenvalue for the Stokes 
equation on $\Omega$, then
\begin{equation}
f(\tilde{k}) \leq C N^{-p} \, . 
\end{equation}
\end{thrm}
\begin{proof}
The proof is rather tedious and technical and
follows essentially from combining the theorem 4.3 in~\cite{zhao2015robust}, 
and theorem 8.1 in~\cite{bornemann2010numerical}.
\end{proof}
The theorem above relies on the observation that
the determinant of a Nystr\"{o}m discretized linear system
is a good approximation of the Fredholm determinant of trace-class
operators, and is an especially good approximation
when the Fredholm determinant is $0$.

\begin{remark}
The same analysis does not carry through for the operator
$\cI - 2 \cD_{k} - 2i\eta \cS_{k} - 2\cW$, 
since it is not a trace class operator.
For brevity, let $\cM_{k} = -2\cD_{k} - 2i\eta \cS_{k} - 2\cW$.
The operator $\cM$ is in $\cJ_{2}(\bL^{2}(\Gamma))$ where
$\cJ_{2}(\bL^{2}(\Gamma))$ is the space of Hilbert-Schmidt operators
on $\bL^{2}(\Gamma)$ (the singular values of the operator are square
summable, as opposed to being summable).
Thus the Fredholm determinant of $\cI + \cM$
is not necessarily convergent. 
However, as noted in~\cite{zhao2015robust}, the convergence
result~\cref{thm:mainconvfreddet}
should be true up to a logarithmic factor in the rate
of convergence, since the singular values of the operator
$\cM_{k}$ decay like $\frac{1}{n}$, and the
Fredholm determinant divergers logarithmically. 
Let $M_{N,k}$ denote the Nystr\"{o}m discretized operator 
corresponding to $M$, where $N$ is the number of discretization points
on the boundary $\Gamma$.
One way of proving such a result would be to construct a definition 
of a determinant for Hilbert-Schmidt operators defined by 
Hilbert~\cite{hilbertbook} and Carleman~\cite{carlemanbook}
as opposed to the Fredholm determinant
and prove analogous results for this case.
Instead, in~\cref{subsec:convannulus}, we demonstrate that
that $\text{det} (I - M^{N,k}) \to 0$ as $N \to \infty$ if and only 
if $k$ is a Dirichlet eigenvalue for the Stokes operator, and that
the convergence to $0$ is exponential at the eigenvalues even in 
this case.
\end{remark}

>>>>

Let $\cJ_{1}(X)$ denote the space of trace class operators 
on $X$, where $X$ is a Hilbert space.
$\cJ_{1}$ is a subspace of compact operators on $X$.
Suppose that $A$ is a compact operator, with eigenvalues
$\lambda_{i}, i=1,2,\ldots$.
Then the operator $A \in \cJ_{1}(X)$ if
$\sum_{i} |\lambda_{i}| < \infty$.
If $A$, is a trace class operator, then we can define 
the Fredholm determinant of the operator $\cI + A$ as
\begin{equation}
\text{det}(\cI +A) = \prod_{i=1}^{\infty} (1+\lambda_{i}) \, .
\end{equation}
It follows from standard results in complex analysis, 
that the Fredholm determinant is well-defined if 
$\text{det}(\cI+A) < \infty$ if $\sum_{i} |\lambda_{i}| < \infty$,
i.e. if $A$ is in trace class.

The following lemma shows that the operator 
$-2\cD_{k} - 2\cW$ is a trace class operator.
\begin{lemma} 
$-2\cD_{k} - 2\cW$ is a trace class operator for all $k$ 
\end{lemma}
The Fredholm determinant of a second-kind Fredholm operator
is another way of detecting when the operator is not invertible. 
The following result summarizes the result. 
\begin{lemma}
If $A$ is a trace-class operator, then
\end{lemma}
